Ignition of Fuel Beds by Firebrands

### 2022-05-14 litreview.tex
Updates to multi-heater study

    The general mechanism for firebrand ignition is considered to be a three step process which starts within the fire itself, often at a significant distance from the ignition point. First, a firebrand is generated as burning material breaks off or is lofted into the air. Second, the firebrand is then lofted through the air and transported to the ignition site. Third, a firebrand lands on a recipient fuel bed and then ignition occurs if the criteria for ignition is met~\cite{Babrauskas2003}. When considering the possibility of ignition during a fire event all three of processes must be considered as each process significantly influences the properties of an ember landing on a fuel bed. For example, the number, energy content, and size of embers that may land on a fuel bed is significantly influence by the surrounding foliage~\cite{Hudson2020EffectsScale} and moisture content of the foliage~\cite{Adusumilli2021FirebrandShrub} which introduces significant variability in ignition sources a fuel bed may encounter. Similarly, the winds transporting a firebrand significantly influence the path of travel and time of flight which influences the energy content~\cite{Sardoy2007, MatvienkoOVA2016} and localized winds around structures may lead to firebrand accumulation~\cite{Suzuki2020a, Suzuki2017, Suzuki2015}. Unfortunately, the generation and ignition processes are significantly less understood than the transport processes~\cite{Manzello2020}. The work contained in this dissertation is focused on the ignition process of the fuel bed by a firebrand and does not consider the generation of firebrands. For further information on firebrand generation processes the reader is referred to review papers considering this process by Fernandez-Pello~\cite{Fernandez-Pello2017} and by Manzello et al.~\cite{Manzello2020}.

    Ignition of a fuel by firebrands has been identified as a source of house loss f
    Developing an increased understanding of the ignition of a fuel bed has been identified as a necessary step to better protect structures from firebrand ignition since at  ~\cite{Waterman1969ExperiemntalGeneration}. but research was limited until the mid 2000's.

    Early research and a significant portion of current research has been conducted by dropping combusting firebrands (either flaming or glowing) or hot metal particles fuel beds in varying configurations that simulate conditions in a wildfire and observing ignition or no ignition outcomes. These works have identified some parameters that appear to control ignition but models created from these studies only provide qualitative predictions of ignition. The lack of models of quantitative accuracy was suggested by Finney et al. in 2013 who identified the lack of a fundamental understanding of what processes occur and how they interact as a primary gap in the knowledge of ignition~\cite{Finney2013}. A similar need was identified by Manzello et al. in 2018~\cite{Manzello2018} and again in 2020~\cite{Manzello2020}. As fires increase in severity, and the WUI continues to expand the number of homes lost to wildfires is likely to significantly increase in the near future. Thus, the need for accurate predictions of ignition has never been higher.

@@

\section{Summary}

>>>>

    The general mechanism for firebrand ignition is considered to be a three step process which starts within the fire itself, often at a significant distance from the ignition point. First, a firebrand is generated as burning material breaks off or is lofted into the air. Second, the firebrand is then lofted through the air and transported to the ignition site. Third, a firebrand lands on a recipient fuel bed and then ignition occurs if the criteria for ignition is met~\cite{Babrauskas2003}. When considering the possibility of ignition during a fire event all three of processes must be considered as each process significantly influences the properties of an ember landing on a fuel bed. For example, the number, energy content, and size of embers that may land on a fuel bed is significantly influence by the surrounding foliage~\cite{Hudson2020EffectsScale} and moisture content of the foliage~\cite{Adusumilli2021FirebrandShrub} which introduces significant variability in ignition sources a fuel bed may encounter. Similarly, the winds transporting a firebrand significantly influence the path of travel and time of flight which influences the energy content~\cite{Sardoy2007, MatvienkoOVA2016} and localized winds around structures may lead to firebrand accumulation~\cite{Suzuki2020a, Suzuki2017, Suzuki2015}. Unfortunately, the generation and ignition processes are significantly less understood than the transport processes~\cite{Manzello2020}. The work contained in this dissertation is focused on the ignition process of the fuel bed by a firebrand and does not consider the generation of firebrands. For further information on firebrand generation processes the reader is referred to review papers considering this process by Fernandez-Pello~\cite{Fernandez-Pello2017}, Manzello et al.~\cite{Manzello2020} and Babrauskas~\cite{Babrauskas2003}.

    Firebrands have been identified as a source of structure loss as early as the 1600s~\cite{Suzuki2020} however studies of firebrand ignition have only begun recently. 
    Perhaps one of the first studies published to specifically address firebrands within the scope of structure loss was conducted by Waterman in 1969~\cite{Waterman1969ExperiemntalGeneration}, however research remained sparse until the early 2000s. Early research and a significant portion of current research has been conducted by dropping combusting firebrands (either flaming or glowing) or hot metal particles fuel beds in varying configurations that simulate conditions in a wildfire and observing ignition or no ignition outcomes. These works have identified some parameters that appear to control ignition but models created from these studies only provide qualitative predictions of ignition. The lack of models of quantitative accuracy was suggested by Finney et al. in 2013 who identified the lack of a fundamental understanding of what processes occur and how they interact as a primary gap in the knowledge of ignition~\cite{Finney2013}. A similar need was identified by Manzello et al. in 2018~\cite{Manzello2018} and again in 2020~\cite{Manzello2020}. As fires increase in severity, and the WUI continues to expand the number of homes lost to wildfires is likely to significantly increase in the near future. Thus, the need for accurate predictions of ignition has never been higher.

@@

### 2022-05-14 manuscript3.tex
Updates to multi-heater study

\section{Abstract}
\label{sec:abtract3}
    The increasing severity of wildfires and the expansion of the wildland urban interface have placed a greater number of homes at risk for being destroyed by wildfires. Ignition of fuel beds on or near homes by embers is a significant source of fire spread and structure loss. Fire prevention and suppression efforts lack an expedient method for estimating the likelihood of ignition for the wide variety of materials around homes.

\section{Introduction}\addvspace{10pt}
\label{sec:introduction3}
    Structure losses due to wildfires have increased markedly in recent years, with over 60\% of homes lost in the last 15 years occurring since 2017~\cite{Barrett2020}. One mechanism that leads to home loss during wildfires is ignition of flammable materials around a home (e.g., pine needles accumulated on a roof) by an ember~\cite{Syphard2019, Suzuki2020, Koo2010a}. The combustion of the flammable materials then ignite the structure itself, resulting the loss of the structure. When seeking to understand the processes that occur leading up to the ignition of a structure by firebrands three processes must be considered: (1) the generation of embers in the fire itself, (2) the transport of the embers to the fuel bed, and (3) the ignition of a fuel bed after the ember lands. This work focuses on event (3), specifically the processes of heat transfer and subsequent ignition that occur after an ember lands on a fuel bed.

    The phenomena that occur after an ember lands can be further broken down into three categories. First the properties of the ember that lands, second the heat transfer between the ember, fuel bed, and the surroundings, and third how the fuel bed responds to the heat transfer from the ember. The response of the fuel bed to the heat transfer from the the ember is influenced by the morphology and chemical composition of the particles in the fuel bed. Fuel bed materials present in wildfires vary significantly in both morphology and chemical composition and knowledge of the effects of each on ignition are qualitative, at best. A previous study evaluated the effects of particle morphology~\cite{Bean2021} and this work seeks to identify how differences in the chemical composition of a fuel bed under attack by firebrands influence the probability of flaming ignition.

    Fuel beds at risk for ignition by embers vary significantly in chemical composition which may result in differences in thermal properties and the chemical species produces during pyrolysis. Works considering ignition of fuel beds with varying concentrations of constituents have found significant variation in ignition characteristics. Cellulose fuel beds have been observed to have a lower ignition threshold when compared to grass and pine needles when put in contact by hot metal particles~\cite{Urban2018}. Interestingly, cellulose undergoes pyrolysis at either higher temperatures or later in the heating process than both lignin and hemicellulose~\cite{Yang2007a, Shotorban2018}. This suggests that the lower ignition temperature of cellulose is not due to pyrolysis occurring at lower temperatures, but to the flammability of pyrolyzates produced.

    One method to evaluate the effects of chemical composition on the flammability of pyrolyzates produced, and thus ignition propensity, is through modeling of the pyrolysis process. Efforts to characterize the pyrolysis of biomass have continually improved the understanding of the chemical processes that occur in both the solid and gas phase of pyrolysis products. Recently, more inclusive pyrolysis models have been developed that improve on the initial single step Arrhenius reactions~\cite{DIBLASI199371} and more closely approximate the reactions occurring during combustion of biomass~\cite{Ranzi2008, Debiagi2015, Dhahak2019} based on the initial chemical composition of the fuels. With the development of models that are applicable to a wide range of chemical compositions and environmental factors the possibility of modelling pyrolysis and ignition of fuel beds is possible. Coupling these models with databases, such as the Bioenergy Feedstock Library~\cite{feedstock} (which contains the chemical composition information for a wide variety of materials that may be vulnerable to ember attack during a wildfire), predictions of ignition potential may be possible without data from ignition experiments. While currently available tools, like combustion models and composition databases, may be useful individually for understanding fuel bed ignition and pyrolysis, a framework linking the individual resources and knowledge to provide a general model of fuel bed ignition is lacking. Unfortunately, the current level of knowledge on the ignition threshold of materials of variable composition does not facilitate such a framework.

@@

\section{Background}
\label{sec:background3}
    Studies considering ignition have primarily focused on the influence of heat transfer and fuel bed conditions when identifying ignition parameters and the observed sensitivities to fuel bed properties are attributed to both heat transfer and chemistry related processes. The probability of ignition generally increases as the energy content of an ember increases~\cite{Hadden2011}. However, further research has indicated that energy content alone is not a sufficient indicator of ignition as a minimum ember temperature (a driver of heat flux rates) is also necessary~\cite{Zak2014}. Additionally, heat loss from embers to the surroundings has a significant effect on ignition~\cite{Fernandez-Pello2015}. While the aforementioned efforts have focused primarily on the impact of ember properties on ignition, efforts focused on the properties of the fuel bed have shown sensitivities to fuel bed material type and particle size~\cite{Urban2018}. Additionally, for fuel beds of constant particle size, an increase in the packing density of the fuel bed resulted in a decrease in the critical heat flux needed for ignition~\cite{Mindykowski2011, Hernandez2017, Rivera2020}. The observed sensitivities to fuel bed properties are likely caused by both heat transfer and chemistry related processes. Changes in the particle size of a fixed material may impact the contact area (e.g., a ball resting on fine sawdust compared to a pile of pine needles) between the ember and fuel bed; potentially reducing the amount of heat transferred through conduction as the contact area decreases. Additionally, when considering a fixed particle size fuel bed, changes in chemical composition may change the heat transfer and heating rate of the materials due to changes in the thermal conductivity of the material affecting the ignition properties of the material. For these reasons, knowledge of the effects of chemical composition on ignition, as it relates to thermal and chemical processes, must be addressed to accurately represent ignition properties for the wide range of materials found in fires resulting in home loss.

@@

    The 50\% probability of flaming ignition with respect to the surface temperature of a resistance heater was determined for five different materials in quiescent conditions and at a wind speed of 5.8\si{\meter\per\second}. The fuel bed materials used were Douglas-fir wood, Douglas-fir bark, pine wood, pine bark, wheat straw, and oak wood. These materials were chosen to represent a range of chemical compositions, and are materials that may be subject to ember attack in a wildland urban interface fire. Table \ref{tab:composition} shows the estimated chemical composition of the materials based on the Bioenergy Feedstock Library~\cite{feedstock}. To create fuel beds of consistent particle size,  materials were granulated and then sorted such that the particles fit through a screen with openings of 2.1\si{\milli\meter} but not through openings of 0.85\si{\milli\meter}. Once sorted, the materials were oven dried at 103\si{\celsius}.

@@

    Two experimental apparatus were used to conduct ignition experiments. The experimental apparatus used for the tests in quiescent conditions is shown in Figure~\ref{fig:speciesApparatus}. The quiescent apparatus used a lever arm to hold the cartridge heater in a fixed position the duration of the test. The heater was inserted into the fuel bed approximately half of the heater diameter (3\si{\milli\meter}). Tests conducted with a wind speed of \si{\meter\per\second} were performed using the apparatus shown in Figure~\ref{fig:speciesWindTunnel}. This apparatus was operated inside of a wind tunnel where the heater was lowered using an automated lowering device that maintained a constant pressure equivalent to that of a 10\si{\gram} ember landing on the fuel bed. The pressure was monitored and adjusted using a PID control system with force measured by a load cell. For all tests the cartridge heater (firebrand surrogate) used was a 50\si{\milli\meter} long 6.35\si{\milli\meter} diameter cartridge heater.

@@

    The temperature of the heater was recorded with a type-K thermocouple for the duration of each test. The temperature of the heater for ranged from 250\si{\celsius} to 750\si{\celsius} and was controlled using PID control logic implemented in LabVIEW. The PID controller kept the heater temperature within 6\% of the setpoint for the duration of the tests. The temperature setpoints of the resistance heater were determined using the three-phase optimal design procedure to efficiently determine the desired ignition probability for the number of tests conducted~\cite{Wu2014}. For tests where flaming ignition did not occur data was recorded for 3000\si{\second} or until the reaction front of the smoldering material reached the edge of the test container. The fixed position of the cartridge heater and the use of the cartridge heater was implemented to maintain consistency between tests of the same material and between test series of different materials. The use of the cartridge heater and the fixed position is not necessarily representative of a burning ember landing on the fuel bed but the consistency of surface temperatures, ability to record temperature, and consistent contact with the fuel bed are essential for comparing results across test series and materials. For the quiescent tests flaming ignition was detected with the use of a BPX 65 photodiode sampled at 1 kHz. If the intensity detected rose above a set threshold flaming ignition was said to occur. For the wind tunnel tests flaming ignition was determined from the rising temperature of the cartridge heater in the presence of flame. Visual detection of flames was also used in cohort with the photodiode measurements. The 50\% probability of ignition was determined from a logistic regression on the outcomes of 25 ignition tests conducted for each of the materials for which flaming ignition was observed. The scikit-learn python package~\cite{scikit-learn} was used to perform the logistic regressions used to predict ignition probabilities.

@@

    The 50\% ignition probability results for the 5.8\si{\meter\per\second} wind speed tests are shown in Table~\ref{tab:composition50temp} along with the results for the 0.1\si{\meter\per\second} tests. From these results there are three observations of note. First, the increase of wind lowered the threshold for ignition probability for the Douglas-fir wood and Pine wood by approximately 30\%. Second, in contrast to the Douglas-fir and Pine wood results, an increase in wind resulted in a decrease in a higher threshold for ignition.  Third, the materials that were not observed to ignite in the quiescent tests were also not observed to ignite in the presence of wind. While the trends for douglas-fir wood and pine wood match those presented in Chapter~\ref{part:manuscript2} and trends reported in other studies of pine needle ignition~\cite{Wang2017} and eucalyptus bark~\cite{Ellis2011, Ellis2015}. The decreasing ignition probability of the oak wood contradicts those observed in the aforementioned studies and this study but is not unprecedented. Ignition studies of various litter layers types ignited by various single firebrands observed a sensitivity to ember location where embers that landed on top of a fuel bed in windy conditions were less likely to cause ignition than those embedded where ignition was enhanced~\cite{Plucinski2008}. This trend was attributed to increased heat loss due to wind. In this study, however, the location of the cartridge heater was consistent across all materials, the energy is not a function of the wind speed, and the flow field around the embers is consistent.  
    \begin{table*}[hpbt]
        \caption{Heater temperature required for 50\% probability of ignition for the materials and conditions tested.}
        \centering
        \begin{tabular}{crrr}
            % \rowcolor{gray!50}
            Material & 0.1 \si{\meter\per\second}(\si{\celsius}) & 5.8 \si{\meter\per\second} (\si{\celsius}) & $\Delta$T (\si{\celsius})\\
            \hline
            Douglas-fir Wood & 574  & 391 & -185 \\
            Pine Wood        & 634  & 430 & -204\\
            Oak Wood         & 663  & \textgreater750 & 87\\
            Wheat Straw      & \textgreater750 & \textgreater750 & -\\
            Pine Bark        & \textgreater750 & \textgreater750 & -\\
            Douglas-Fir Bark & \textgreater750 & \textgreater750 & -
        \end{tabular}
        \label{tab:composition50temp}
    \end{table*}

@@

    Perhaps unsurprisingly, Douglas-fir bark and pine bark, with the highest lignin content, were not observed to ignite for the conditions tested. This suggests that materials with high lignin contents may be the least likely to ignite when exposed to embers and thus are the safest with respect to flaming ignition risk around homes. This does not, however, account for the potential for smoldering ignition and a subsequent smoldering to flaming transition that may occur. Thermogravametric analysis experiments have shown that while lignin begins to decompose at the lowest temperature of the three primary components, it decomposes at the slowest rate and at a much wider range of temperatures~\cite{Yang2007a}. It was also observed that the peak gas production of \ce{CO}, \ce{CO2}, and \ce{CH4} occurs at a higher temperatures than hemicellulose and cellulose. In contrast to Douglas-fir bark, wheat straw had the lowest estimated lignin concentration but was also not observed to ignite. When comparing wheat straw to Douglas-fir wood the higher lignin concentration suggests that Douglas-fir wood is likely to have a higher ignition temperature than wheat straw. This discrepancy is attributed to two material attributes. First, the pyrolysis process is complex and there is not a clear linear relationship between the composition of the fuel and flaming ignition. It appears that materials of different compositions are capable of producing gaseous products of near equal ignitability but the proportion of each constituent for which ignition to occurs most readily is unclear. Second, the thermal conductivity of the materials likely varies significantly which impacts the temperature gradients and mass of the material above the pyrolysis temperature, further obfuscating the effect of composition on ignition.

>>>>

\section{Abstract}
\label{sec:abtract3}
    The increasing severity of wildfires and the expansion of the wildland urban interface have placed a greater number of homes at risk for being destroyed by wildfires. Ignition of fuel beds on or near homes by firebrands is a significant source of fire spread and structure loss. Fire prevention and suppression efforts lack an expedient method for estimating the likelihood of ignition for the wide variety of materials around homes.

\section{Introduction}\addvspace{10pt}
\label{sec:introduction3}

    The phenomena that occur after an firebrand lands can be further broken down into three categories. First the properties of the firebrand that lands, second the heat transfer between the firebrand, fuel bed, and the surroundings, and third how the fuel bed responds to the heat transfer from the firebrand. The response of the fuel bed to the heat transfer from the the firebrand is influenced by the morphology and chemical composition of the particles in the fuel bed. Fuel bed materials present in wildfires vary significantly in both morphology and chemical composition and knowledge of the effects of each on ignition are qualitative, at best. A previous study evaluated the effects of particle morphology~\cite{Bean2021} and this work seeks to identify how differences in the chemical composition of a fuel bed under attack by firebrands influence the probability of flaming ignition.

    Fuel beds at risk for ignition by firebrands vary significantly in chemical composition which may result in differences in thermal properties and the chemical species produces during pyrolysis. Works considering ignition of fuel beds with varying concentrations of constituents have found significant variation in ignition characteristics. Cellulose fuel beds have been observed to have a lower ignition threshold when compared to grass and pine needles when put in contact by hot metal particles~\cite{Urban2018}. Interestingly, cellulose undergoes pyrolysis at either higher temperatures or later in the heating process than both lignin and hemicellulose~\cite{Yang2007a, Shotorban2018}. This suggests that the lower ignition temperature of cellulose is not due to pyrolysis occurring at lower temperatures, but to the flammability of pyrolyzates produced.

    One method to evaluate the effects of chemical composition on the flammability of pyrolyzates produced, and thus ignition propensity, is through modeling of the pyrolysis process. Efforts to characterize the pyrolysis of biomass have continually improved the understanding of the chemical processes that occur in both the solid and gas phase of pyrolysis products. Recently, more inclusive pyrolysis models have been developed that improve on the initial single step Arrhenius reactions~\cite{DIBLASI199371} and more closely approximate the reactions occurring during combustion of biomass~\cite{Ranzi2008, Debiagi2015, Dhahak2019} based on the initial chemical composition of the fuels. With the development of models that are applicable to a wide range of chemical compositions and environmental factors the possibility of modelling pyrolysis and ignition of fuel beds is possible. Coupling these models with databases, such as the Bioenergy Feedstock Library~\cite{feedstock} (which contains the chemical composition information for a wide variety of materials that may be vulnerable to firebrand attack during a wildfire), predictions of ignition potential may be possible without data from ignition experiments. While currently available tools, like combustion models and composition databases, may be useful individually for understanding fuel bed ignition and pyrolysis, a framework linking the individual resources and knowledge to provide a general model of fuel bed ignition is lacking. Unfortunately, the current level of knowledge on the ignition threshold of materials of variable composition does not facilitate such a framework.

    Studies considering ignition have primarily focused on the influence of heat transfer and fuel bed conditions when identifying ignition parameters and the observed sensitivities to fuel bed properties are attributed to both heat transfer and chemistry related processes. The probability of ignition generally increases as the energy content of an firebrand increases~\cite{Hadden2011}. However, further research has indicated that energy content alone is not a sufficient indicator of ignition as a minimum firebrand temperature (a driver of heat flux rates) is also necessary~\cite{Zak2014}. Additionally, heat loss from firebrands to the surroundings has a significant effect on ignition~\cite{Fernandez-Pello2015}. While the aforementioned efforts have focused primarily on the impact of firebrand properties on ignition, efforts focused on the properties of the fuel bed have shown sensitivities to fuel bed material type and particle size~\cite{Urban2018}. Additionally, for fuel beds of constant particle size, an increase in the packing density of the fuel bed resulted in a decrease in the critical heat flux needed for ignition~\cite{Mindykowski2011, Hernandez2017, Rivera2020}. The observed sensitivities to fuel bed properties are likely caused by both heat transfer and chemistry related processes. Changes in the particle size of a fixed material may impact the contact area (e.g., a ball resting on fine sawdust compared to a pile of pine needles) between the firebrand and fuel bed; potentially reducing the amount of heat transferred through conduction as the contact area decreases. Additionally, when considering a fixed particle size fuel bed, changes in chemical composition may change the heat transfer and heating rate of the materials due to changes in the thermal conductivity of the material affecting the ignition properties of the material. For these reasons, knowledge of the effects of chemical composition on ignition, as it relates to thermal and chemical processes, must be addressed to accurately represent ignition properties for the wide range of materials found in fires resulting in home loss.

@@

@@

    The 50\% probability of flaming ignition with respect to the surface temperature of a resistance heater was determined for five different materials in quiescent conditions and at a wind speed of 5.8\si{\meter\per\second}. The fuel bed materials used were Douglas-fir wood, Douglas-fir bark, pine wood, pine bark, wheat straw, and oak wood. These materials were chosen to represent a range of chemical compositions, and are materials that may be subject to firebrand attack in a wildland urban interface fire. Table \ref{tab:composition} shows the estimated chemical composition of the materials based on the Bioenergy Feedstock Library~\cite{feedstock}. To create fuel beds of consistent particle size,  materials were granulated and then sorted such that the particles fit through a screen with openings of 2.1\si{\milli\meter} but not through openings of 0.85\si{\milli\meter}. Once sorted, the materials were oven dried at 103\si{\celsius}.

@@

    Two experimental apparatus were used to conduct ignition experiments. The experimental apparatus used for the tests in quiescent conditions is shown in Figure~\ref{fig:speciesApparatus}. The quiescent apparatus used a lever arm to hold the cartridge heater in a fixed position the duration of the test. The heater was inserted into the fuel bed approximately half of the heater diameter (3\si{\milli\meter}). Tests conducted with a wind speed of \si{\meter\per\second} were performed using the apparatus shown in Figure~\ref{fig:speciesWindTunnel}. This apparatus was operated inside of a wind tunnel where the heater was lowered using an automated lowering device that maintained a constant pressure equivalent to that of a 10\si{\gram} firebrand landing on the fuel bed. The pressure was monitored and adjusted using a PID control system with force measured by a load cell. For all tests the cartridge heater (firebrand surrogate) used was a 50\si{\milli\meter} long 6.35\si{\milli\meter} diameter cartridge heater.

@@

    The temperature of the heater was recorded with a type-K thermocouple for the duration of each test. The temperature of the heater for ranged from 250\si{\celsius} to 750\si{\celsius} and was controlled using PID control logic implemented in LabVIEW. The PID controller kept the heater temperature within 6\% of the setpoint for the duration of the tests. The temperature setpoints of the resistance heater were determined using the three-phase optimal design procedure to efficiently determine the desired ignition probability for the number of tests conducted~\cite{Wu2014}. For tests where flaming ignition did not occur data was recorded for 3000\si{\second} or until the reaction front of the smoldering material reached the edge of the test container. The fixed position of the cartridge heater and the use of the cartridge heater was implemented to maintain consistency between tests of the same material and between test series of different materials. The use of the cartridge heater and the fixed position is not necessarily representative of a burning firebrand landing on the fuel bed but the consistency of surface temperatures, ability to record temperature, and consistent contact with the fuel bed are essential for comparing results across test series and materials. For the quiescent tests flaming ignition was detected with the use of a BPX 65 photodiode sampled at 1 kHz. If the intensity detected rose above a set threshold flaming ignition was said to occur. For the wind tunnel tests flaming ignition was determined from the rising temperature of the cartridge heater in the presence of flame. Visual detection of flames was also used in cohort with the photodiode measurements. The 50\% probability of ignition was determined from a logistic regression on the outcomes of 25 ignition tests conducted for each of the materials for which flaming ignition was observed. The scikit-learn python package~\cite{scikit-learn} was used to perform the logistic regressions used to predict ignition probabilities.

@@

    The 50\% ignition probability results for the 5.8\si{\meter\per\second} wind speed tests are shown in Table~\ref{tab:composition50temp} along with the results for the 0.1\si{\meter\per\second} tests. From these results there are three observations of note. First, the increase of wind lowered the threshold for ignition probability for the Douglas-fir wood and Pine wood by approximately 30\%. Second, in contrast to the Douglas-fir and Pine wood results, an increase in wind resulted in a decrease in a higher threshold for ignition.  Third, the materials that were not observed to ignite in the quiescent tests were also not observed to ignite in the presence of wind. While the trends for douglas-fir wood and pine wood match those presented in Chapter~\ref{part:manuscript2} and trends reported in other studies of pine needle ignition~\cite{Wang2017} and eucalyptus bark~\cite{Ellis2011, Ellis2015}. The decreasing ignition probability of the oak wood contradicts those observed in the aforementioned studies and this study but is not unprecedented. Ignition studies of various litter layers types ignited by various single firebrands observed a sensitivity to firebrand location where firebrands that landed on top of a fuel bed in windy conditions were less likely to cause ignition than those embedded where ignition was enhanced~\cite{Plucinski2008}. This trend was attributed to increased heat loss due to wind. In this study, however, the location of the cartridge heater was consistent across all materials, the energy is not a function of the wind speed, and the flow field around the firebrands is consistent.  
    \begin{table*}[hpbt]
        \caption{Heater temperature required for 50\% probability of ignition for the materials and conditions tested.}
        \centering
        \begin{tabular}{crrr}
            % \rowcolor{gray!50}
            Material & 0.1 \si{\meter\per\second}(\si{\celsius}) & 5.8 \si{\meter\per\second} (\si{\celsius}) & $\Delta$T (\si{\celsius})\\
            \hline
            Douglas-fir Wood & 574  & 391 & -185 \\
            Pine Wood        & 634  & 430 & -204\\
            Oak Wood         & 663  & \textgreater750 & 87\\
            Wheat Straw      & \textgreater750 & \textgreater750 & -\\
            Pine Bark        & \textgreater750 & \textgreater750 & -\\
            Douglas-Fir Bark & \textgreater750 & \textgreater750 & -
        \end{tabular}
        \label{tab:composition50temp}
    \end{table*}

@@

    Perhaps unsurprisingly, Douglas-fir bark and pine bark, with the highest lignin content, were not observed to ignite for the conditions tested. This suggests that materials with high lignin contents may be the least likely to ignite when exposed to firebrands and thus are the safest with respect to flaming ignition risk around homes. This does not, however, account for the potential for smoldering ignition and a subsequent smoldering to flaming transition that may occur. Thermogravametric analysis experiments have shown that while lignin begins to decompose at the lowest temperature of the three primary components, it decomposes at the slowest rate and at a much wider range of temperatures~\cite{Yang2007a}. It was also observed that the peak gas production of \ce{CO}, \ce{CO2}, and \ce{CH4} occurs at a higher temperatures than hemicellulose and cellulose. In contrast to Douglas-fir bark, wheat straw had the lowest estimated lignin concentration but was also not observed to ignite. When comparing wheat straw to Douglas-fir wood the higher lignin concentration suggests that Douglas-fir wood is likely to have a higher ignition temperature than wheat straw. This discrepancy is attributed to two material attributes. First, the pyrolysis process is complex and there is not a clear linear relationship between the composition of the fuel and flaming ignition. It appears that materials of different compositions are capable of producing gaseous products of near equal ignitability but the proportion of each constituent for which ignition to occurs most readily is unclear. Second, the thermal conductivity of the materials likely varies significantly which impacts the temperature gradients and mass of the material above the pyrolysis temperature, further obfuscating the effect of composition on ignition.
